---
title: "Feature Selection"
output: html_document
date: "2025-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Feature Selection Methods going to use: Information Gain, Mutual Information, Chi-Squared, Fisher's Score, Pearson's Correlation Coefficient (R), Variance Threshold
#Start with 20 features maybe eliminate down to like top 10 then once have that can design experiment

##Information Gain packages loaded in
```{r}
install.packages("funModeling")
library(funModeling)
```

##Load in data
```{r}
data <- read.csv("~/Desktop/Personal/Football ML/Football Data/Punt Data/PuntDataPulled.csv")
```

##Run Information Gain Ratio, omit height and weight

```{r}
IG_PlayerIDP <- gain_ratio(data$PlayerIDP, data$Efficiency)
IG_PLocID <- gain_ratio(data$PLocID, data$Efficiency)
IG_PlayerIDLS <- gain_ratio(data$PlayerIDLS, data$Efficiency)
IG_SnapLocID <- gain_ratio(data$SnapLocID, data$Efficiency)
IG_PDate <- gain_ratio(data$PDate, data$Efficiency)
IG_Snaptime <- gain_ratio(data$Snaptime, data$Efficiency)
IG_OP <- gain_ratio(data$OP, data$Efficiency)
IG_H2F <- gain_ratio(data$H2F, data$Efficiency)
IG_Hang <- gain_ratio(data$Hang, data$Efficiency)
IG_Distance <- gain_ratio(data$Distance, data$Efficiency)
IG_Practice <- gain_ratio(data$Practice, data$Efficiency)
IG_Game <- gain_ratio(data$Game, data$Efficiency)
IG_Precipitation <- gain_ratio(data$precipitation, data$Efficiency)
IG_Wind <- gain_ratio(data$Wind, data$Efficiency)
IG_Grass <- gain_ratio(data$Grass, data$Efficiency)
IG_Turf <- gain_ratio(data$Turf, data$Efficiency)
IG_Temp <- gain_ratio(data$Temp, data$Efficiency)
```

##Mutual Information (MI) loading packages in
```{r}
install.packages("FSelectorRcpp")
library(FSelectorRcpp)
```
##Running MI on only discrete data, omit height and weight as well
```{r}
MI_PlayerIDP <- mutinformation(data$PlayerIDP, data$Efficiency, method = "emp")
MI_PlocID <- mutinformation(data$PLocID, data$Efficiency, method = "emp")
MI_PlayerIDLS <- mutinformation(data$PlayerIDLS, data$Efficiency, method = "emp")
MI_SnapLocID <- mutinformation(data$SnapLocID, data$Efficiency, method = "emp")
MI_PDate <- mutinformation(data$PDate, data$Efficiency, method = "emp")
#mutinformation(data$Snaptime, data$Efficiency, method = "emp") ##ommitted due to continuous
#mutinformation(data$OP, data$Efficiency, method = "emp")       ##ommitted due to continuous
#mutinformation(data$H2F, data$Efficiency, method = "emp")     ##ommitted due to continuous
#mutinformation(data$Hang, data$Efficiency, method = "emp")
MI_Distance <- mutinformation(data$Distance, data$Efficiency, method = "emp")
MI_Practice <- mutinformation(data$Practice, data$Efficiency, method = "emp")
MI_Game <- mutinformation(data$Game, data$Efficiency, method = "emp")
MI_Wind <- mutinformation(data$Wind, data$Efficiency, method = "emp")
MI_Grass <- mutinformation(data$Grass, data$Efficiency, method = "emp")
MI_Turf <- mutinformation(data$Turf, data$Efficiency, method = "emp")
```


