---
title: "Feature Selection"
output: html_document
date: "2025-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Feature Selection Methods going to use: Information Gain, Mutual Information, Chi-Squared, Fisher's Score, Pearson's Correlation Coefficient (R), Variance Threshold
#Start with 20 features maybe eliminate down to like top 10 then once have that can design experiment

##Information Gain packages loaded in
```{r}
install.packages("funModeling")
library(funModeling)
install.packages("FSelector")
library(FSelectorRcpp)
install.packages("CORElearn")
library(CORElearn)
```

##Load in data
```{r}
data <- read.csv("~/Desktop/Personal/Football ML/Football Data/Punt Data/PuntDataPulled.csv")
data_norm <- read.csv("~/Desktop/Personal/Football ML/Football Data/Punt Data/PuntDataFinal.csv")
```

##Run Information Gain Ratio, omit height and weight for hang


```{r}
install.packages("writexl")
library(writexl)
```




```{r}
IG_PlayerIDP <- gain_ratio(data$PlayerIDP, data_norm$Hang)
IG_PLocID <- gain_ratio(data$PLocID, data_norm$Hang)
IG_PlayerIDLS <- gain_ratio(data$PlayerIDLS, data_norm$Hang)
IG_SnapLocID <- gain_ratio(data$SnapLocID, data_norm$Hang)
IG_PDate <- gain_ratio(data$PDate, data_norm$Hang)
IG_Snaptime <- gain_ratio(data$Snaptime, data_norm$Hang)
IG_OP <- gain_ratio(data$OP, data_norm$Hang)
IG_H2F <- gain_ratio(data$H2F, data_norm$Hang)
IG_Distance <- gain_ratio(data$Distance, data_norm$Hang)
IG_Practice <- gain_ratio(data$Practice, data_norm$Hang)
IG_Game <- gain_ratio(data$Game, data_norm$Hang)
IG_Precipitation <- gain_ratio(data$precipitation, data_norm$Hang)
IG_Wind <- gain_ratio(data$Wind, data_norm$Hang)
IG_Grass <- gain_ratio(data$Grass, data_norm$Hang)
IG_Turf <- gain_ratio(data$Turf, data_norm$Hang)
IG_Temp <- gain_ratio(data$Temp, data_norm$Hang)
```

##Mutual Information (MI) loading packages in
```{r}
install.packages("FSelectorRcpp")
library(FSelectorRcpp)
install.packages ("infotheo")
library(infotheo)
```

```{r}
##Creating Bins for Hang for MI
summary(data_norm$Hang)


data_norm$HangCat <- cut (
  data_norm$Hang, 
  breaks = c(-Inf, -0.46696, 0.03192, 0.16007, 0.57809, Inf), 
  labels = c(1,2,3,4,5),
  right = TRUE
)

##Bins for Snaptime
summary(data_norm$Snaptime)


data_norm$SnaptimeCat <- cut (
  data_norm$Hang, 
  breaks = c(-Inf, -0.5547, 0.02934, 0.7174, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)


##Bin forOP
summary(data_norm$OP)


data_norm$OPCat <- cut (
  data_norm$OP, 
  breaks = c(-Inf, -0.50025,.04587, 0.64660, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)



##bin for H2f
summary(data_norm$H2F)


data_norm$H2FCat <- cut (
  data_norm$H2F, 
  breaks = c(-Inf, -0.53356, 0.05513, 0.64383, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)

summary(data_norm$precipitation)


data_norm$precipitationCat <- cut (
  data_norm$precipitation, 
  breaks = c(-Inf, -0.4102,  -1.777, Inf), 
  labels = c(1,2,3),
  right = TRUE
)

summary(data_norm$Temp)


data_norm$TempCat <- cut (
  data_norm$Temp, 
  breaks = c(-Inf, -1.0226, 0.2533, 0.8536, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)


summary(data_norm$PDate)


data_norm$HPDateCat <- cut (
  data_norm$PDate, 
  breaks = c(-Inf, -0.9405, -0.1511, 0.9114, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)

summary()

summary(data_norm$Distance)


data_norm$DistanceCat <- cut (
  data_norm$Distance, 
  breaks = c(-Inf, -0.455240, -0.041008, 0.511302, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)


summary(data_norm$Wind)


data_norm$WindCat <- cut (
  data_norm$Wind, 
  breaks = c(-Inf, -0.8972, 0.4706, 0.7444, Inf), 
  labels = c(1,2,3,4),
  right = TRUE
)

```

##Running MI on only discrete data, omit height and weight as well
```{r}
MI_PlayerIDP <- mutinformation(data_norm$PlayerIDP, data_norm$HangCat, method = "emp")
MI_PlocID <- mutinformation(data_norm$PLocID, data_norm$HangCat, method = "emp")
MI_PlayerIDLS <- mutinformation(data_norm$PlayerIDLS, data_norm$HangCat, method = "emp")
MI_SnapLocID <- mutinformation(data_norm$SnapLocID, data_norm$HangCat, method = "emp")
MI_PDate <- mutinformation(data$PDate, data_norm$HangCat, method = "emp")
MI_Snaptime <- mutinformation(data_norm$SnaptimeCat, data_norm$HangCat, method = "emp") ##ommitted due to continuous
MI_OP <- mutinformation(data_norm$OPCat, data$Efficiency, method = "emp")       ##ommitted due to continuous
MI_H2F <- mutinformation(data_norm$H2FCat, data$Efficiency, method = "emp")     ##ommitted due to continuous
MI_Distance <- mutinformation(data_norm$DistanceCat, data_norm$HangCat, method = "emp")
MI_Practice <- mutinformation(data_norm$Practice, data_norm$HangCat, method = "emp")
MI_Game <- mutinformation(data_norm$Game, data_norm$HangCat, method = "emp")
MI_Wind <- mutinformation(data_norm$WindCat, data_norm$HangCat, method = "emp")
MI_Grass <- mutinformation(data_norm$Grass, data_norm$HangCat, method = "emp")
MI_Turf <- mutinformation(data_norm$Turf, data_norm$HangCat, method = "emp")
MI_Precipitation <- mutinformation(data_norm$precipitationCat, data_norm$HangCat, method = "emp")
MI_Temp <- mutinformation(data_norm$TempCat, data_norm$HangCat, method = "emp")
```

##Installing Chi-Squared packages
```{r}
install.packages("MASS")
library(MASS)
```



##Creating Contingency Tables
```{r}
##Creates Bins for Distance Feature
summary(data$Distance)
data$DistanceCat <- cut (
  data$Distance, 
  breaks = c(-Inf, 37, 46, Inf),
  labels = c("Short", "Medium", "Long"),
  right = TRUE
)
## Contingency Table for Distance and Effiency
dist = data.frame(data$DistanceCat, data$Hang)
dist = table(data$DistanceCat, data$Hang)
print(dist)

## Game Bins
data$GameCat <- cut (
  data$Game,
  breaks = c(-Inf, 1, 4, Inf),
  labels = c("Practice", "Non-Conference", "Conference"),
  right = TRUE
)

##Contingency Table For Game and Effiency

game = data.frame (data$GameCat, data$Hang)
game = table(data$GameCat, data$Hang)
print(game)

##Grass Bins
data$GrassCat <- cut (
  data$Grass,
  breaks = c(-Inf, 0, 1),
  labels = c("Grass", "Turf"),
  right = TRUE
)

##Grass Contingency Table

grass = data.frame(data$GrassCat, data$Hang)
grass = table(data$GrassCat, data$Hang)
print(grass)

##H2F Bins

data$H2FCat <- cut (
  data$H2F,
  breaks = c(-Inf, 1.220, 1.4, Inf),
  labels = c("Great", "Average", "Poor"),
  right = TRUE
)

##H2F Contingency Table

H2F = data.frame(data$H2FCat, data$Hang)
H2F = table(data$H2FCat, data$Hang)
print(H2F)


##Hang Bins
data$HangCat <- cut (
  data$Hang,
  breaks = c(-Inf, 3.22, 3.9, 4.530),
  labels = c("Poor", "Average", "Great"),
  right = TRUE
)

##Hang Contingency Table
Hang = data.frame(data$HangCat, data$Hang)
Hang = table(data$HangCat, data$Hang)
print(Hang)

##Op Bins
data$OPCat <- cut (
  data$OP,
  breaks = c(-Inf, 1.970, 2.18, 2.680),
  labels = c("Great", "Average", "Poor"),
  right = TRUE
)
##OP Table
OP = data.frame(data$OPCat, data$Hang)
OP = table(data$OPCat, data$Hang)
print(OP)

##PlayerIDLS Bins

data$PlayerIDLSCat <- cut (
  data$PlayerIDLS,
  breaks = c(-Inf, 41, 44, 58, 98),
  labels = c("Gus", "Charlie", "Rob", "Bryce"),
  right = TRUE
)

##PlayerIDLS Table

PlayerIDLS = data.frame(data$PlayerIDLSCat, data$Hang)
PlayerIDLS = table(data$PlayerIDLSCat, data$Hang)
print(PlayerIDLS)

##PlayerIDP
data$PlayerIDPCat <- cut (
  data$PlayerIDP,
  breaks = c(-Inf, 61, 96),
  labels = c("Charlie", "Josh"),
  right = TRUE
)

##PlayerIDP Table

PlayerIDP = data.frame(data$PlayerIDPCat, data$Hang)
PlayerIDP = table(data$PlayerIDPCat, data$Hang)
print(PlayerIDP)

##PLocID Bins

data$PLocIDCat <- cut (
  data$PLocID,
  breaks = c(-Inf, 1, 2, 3),
  labels = c("Poor", "Average", "Great"),
  right = TRUE
)

##PLocID Table

PLocID = data.frame(data$PLocIDCat, data$Hang)
PLocID = table(data$PLocIDCat, data$Hang)
print(PLocID)


##Practice Bins

data$PracticeCat <- cut (
  data$Practice,
  breaks = c(-Inf, 0, 1),
  labels = c("Game", "Practice"),
  right = TRUE
)

##Practice Table

Practice = data.frame(data$PracticeCat, data$Hang)
Practice = table(data$PracticeCat, data$Hang)
print(Practice)

##SnapLocID Bins

data$SnapLocIDCat <- cut (
  data$SnapLocID,
  breaks = c(-Inf, 1, 2, 3),
  labels = c("Poor", "Average", "Great"),
  right = TRUE
  
)

##SnapLocID Table

SnapLocID = data.frame(data$SnapLocIDCat, data$Hang)
SnapLocID = table(data$SnapLocIDCat, data$Hang)
print(SnapLocID)

##Turf Bins

data$TurfCat <- cut (
  data$Practice,
  breaks = c(-Inf, 0, 1),
  labels = c("Grass", "Turf"),
  right = TRUE
)

## Turf Table

Turf = data.frame(data$TurfCat, data$Hang)
Turf = table(data$TurfCat, data$Hang)
print(Turf)

##Wind Bins

data$WindCat <- cut (
  data$Wind,
  breaks = c(-Inf, 10, 15, Inf),
  labels = c("Low", "Average", "High"),
  right = TRUE
)

##Wind Table

Wind = data.frame(data$WindCat, data$Hang)
Wind = table(data$WindCat, data$Hang)
print(Wind)

##Temp Bins

data$TempCat <- cut (
  data$Temp,
  breaks = c(-Inf, 52.79,66.33, Inf),
  labels = c("Cool", "Average", "Warm"),
  right = TRUE
)

Temp = data.frame(data$TempCat, data$Hang)
Temp = table(data$TempCat, data$Hang)
print(Temp)

##Snaptime Bins

data$SnaptimeCat <- cut (
  data$Snaptime,
  breaks = c(-Inf, .7, .78, Inf),
  labels = c("Great", "Average", "Poor"),
  right = TRUE
)

##Snaptime Table

Snaptime = data.frame(data$SnaptimeCat, data$Hang)
Snaptime = table(data$SnaptimeCat, data$Hang)
print(Snaptime)


##Precipitation Bins

data$precipitationCat <- cut (
  data$precipitation,
  breaks = c(-Inf, 0.1, .1764, 2),
  labels = c("Low", "Medium", "High"),
  right = TRUE
)

##Precipitation Table

Precipitation = data.frame(data$precipitationCat, data$Hang)
Precipitation = table(data$precipitationCat, data$Hang)
print(Precipitation)





##Date Conversion

data$PDate <- as.Date(data$PDate)
reference_date <- as.Date("2024-08-01")
data$DaysSinceAug1 <- as.numeric(data$PDate - reference_date)

summary(data$DaysSinceAug1)

##Date Bins

data$DaysSinceAug1Cat <- cut (
  data$DaysSinceAug1,
  breaks = c(-Inf, 29, 51,Inf),
  labels = c("PreSeason", "NonConference", "Conference")
)

DaysSinceAug1 <- data.frame(data$DaysSinceAug1Cat, data$Hang)
DaysSinceAug1 <- table(data$DaysSinceAug1Cat, data$Hang)
print(DaysSinceAug1)


```
##Applying Chi2
```{r}
DistanceChi <- print(chisq.test(dist)) ## test for dist
GameChi <- print(chisq.test(game)) ##test for game
GrassChi <- print(chisq.test(grass)) ##Test for grass
H2FChi <- print(chisq.test(H2F)) ## test for hand to foot
HangChi <- print(chisq.test(Hang)) ## test for Hang
OPChi <- print(chisq.test(OP)) ##test for OP
PlayerIDLSChi <- print(chisq.test(PlayerIDLS)) ##test for PlayerIDLS
PlayerIDPChi <- print(chisq.test(PlayerIDP)) ##test for PlayerIDP
PLocIDChi <- print(chisq.test(PLocID)) ##test for PLocID
PracticeChi <- print(chisq.test(Practice)) ##test for Practice
SnapLocIDChi <- print(chisq.test(SnapLocID)) ##test for SnapLocID
TurfChi <- print(chisq.test(Turf)) ##test for Turf
WindChi <- print(chisq.test(Wind)) ##test for Wind
TempChi <- print(chisq.test(Temp)) ##test for Temp
SnaptimeChi <- print(chisq.test(Snaptime)) ##test for Snaptime
PrecipitationChi <- print(chisq.test(Precipitation)) ##test for Precipitation
DateChi <- print(chisq.test(DaysSinceAug1)) ##Test for Date
```


##Fisher Information Loading
```{r}
numericdata <- data_norm[sapply(data_norm, is.numeric)]
sigmas <- sapply(numericdata, sd, na.rm = TRUE)

fisher_info <- function(x_col, sigma) {
  n <- sum(!is.na(x_col))  # exclude NA values from n
  n / sigma^2
}

fisher_info_values <-mapply(fisher_info, numericdata, sigmas)

print(fisher_info_values)

FIDistance <- fisher_info_values[10]
FIGame <- fisher_info_values[13]
FIGrass <- fisher_info_values[16]
FIH2F <- fisher_info_values[8]
FIOP <- fisher_info_values[7]
FIPDate <- fisher_info_values[5]
FIPlayerIDLS <- fisher_info_values[3]
FIPlayerIDP <- fisher_info_values[1]
FIPLocID <- fisher_info_values[2]
FIPractice <- fisher_info_values[12]
FIPrecipitation <- fisher_info_values[14]
FISnapLocID <- fisher_info_values[4]
FISnaptime <- fisher_info_values[6]
FITemp <- fisher_info_values[18]
FITurf <- fisher_info_values[17]
FIWind <- fisher_info_values[15]
```
                            


```{r}

fisher_score_manual <- function(X, y) {
  classes <- unique(y)
 
  classes <- classes[!is.na(classes)]
  
  if (length(classes) != 2) {
    stop(paste("Only supports binary classification. Found", length(classes), "classes."))
  }
  

  classes <- sort(classes) 
  
  scores <- sapply(X, function(feature) {

    if (!is.numeric(feature)) {
      warning(paste("Skipping non-numeric feature:", deparse(substitute(feature))))
      return(NA) # Return NA for non-numeric features
    }
    
    class1 <- feature[y == classes[1]]
    class2 <- feature[y == classes[2]]
    

    if (length(class1) < 2 || all(is.na(class1)) || var(class1, na.rm = TRUE) == 0) {
      var1 <- 0 # Treat as zero variance if not enough data or constant
    } else {
      var1 <- var(class1, na.rm = TRUE)
    }

    if (length(class2) < 2 || all(is.na(class2)) || var(class2, na.rm = TRUE) == 0) {
      var2 <- 0 # Treat as zero variance if not enough data or constant
    } else {
      var2 <- var(class2, na.rm = TRUE)
    }

    mean_diff <- abs(mean(class1, na.rm = TRUE) - mean(class2, na.rm = TRUE))
    var_sum <- var1 + var2
    
    if (is.na(var_sum) || var_sum == 0) {
      return(0) # Return 0 if variance sum is zero or NA (features offer no discrimination)
    } else {
      return((mean_diff^2) / var_sum)
    }
  })
  

  scores <- scores[!is.na(scores)]
  
  sort(scores, decreasing = TRUE)
}

median_hang <- median(data_norm$Hang, na.rm = TRUE)


data_norm$HangCat <- cut(
  data_norm$Hang,
  breaks = c(-Inf, median_hang, Inf),
  labels = c("Low", "High"), # Meaningful labels for your two classes
  right = TRUE,              # Intervals are (lower, upper]
  include.lowest = TRUE      # Ensures the minimum value is included in the first bin
)


print("Levels of data_norm$HangCat after splitting at median:")
print(levels(data_norm$HangCat))
print(paste("Number of levels:", length(levels(data_norm$HangCat)))) # Should confirm 2 levels


numeric_cols <- names(data_norm)[sapply(data_norm, is.numeric)]
.
X <- data_norm[, setdiff(numeric_cols, "Hang")] 

X <- as.data.frame(X)


y <- data_norm$HangCat

fisher_scores <- fisher_score_manual(X, y)
print("Fisher Scores:")
print(fisher_scores)
```

```{r}
FSDistance <- fisher_scores[2]
FSGame <- fisher_scores[6]
FSGrass <- fisher_scores[12]
FSH2F <- fisher_scores[3]
FSOP <- fisher_scores[4]
FSPDate <- fisher_scores[18]
FSPlayerIDLS <- fisher_scores[17]
FSPlayerIDP <- fisher_scores[7]
FSPLocID <- fisher_scores[8]
FSPractice <- fisher_scores[5]
FSSnapLocID <- fisher_scores[11]
FSSnaptime <- fisher_scores[15]
FSTemp <- fisher_scores[16]
FSTurf <- fisher_scores[12]
FSWind <- fisher_scores[10]
FSPrecipitation <- fisher_scores[14]
```


##Pearson Correlation Coefficient (R)

##Installing COR Package
```{r}
install.packages("COR")
library(COR)
```


##Converting Y (Data_norm$efficiency) from factor to numeric col_type
```{r}
data_norm$Efficiency <- as.numeric(as.character(data_norm$Hang))
str(data_norm$Hang)
```

##PCC Calculation, y = efficiency, x = different features
```{r}
PCC_Distance <- cor.test(data_norm$Distance, data_norm$Hang, method = "pearson")
PCC_Game<- cor.test(data_norm$Game, data_norm$Hang, method = "pearson")
PCC_Grass <- cor.test(data_norm$Grass, data_norm$Hang, method ="pearson")
PCC_H2F <- cor.test(data_norm$H2F, data_norm$Hang, method = "pearson")
PCC_OP <- cor.test(data_norm$OP, data_norm$Hang, method = "pearson")
PCC_PDate <- cor.test(data_norm$PDate, data_norm$Hang, method = "pearson")
PCC_PlayerIDLS <- cor.test(data_norm$PlayerIDLS, data_norm$Hang, method = "pearson")
PCC_PlayerIDP <- cor.test(data_norm$PlayerIDP, data_norm$Hang, method = "pearson")
PCC_PLocID <- cor.test(data_norm$PLocID, data_norm$Hang, method = "pearson")
PCC_Practice <- cor.test(data_norm$Practice, data_norm$Hang, method = "pearson")
PCC_Precipitation<- cor.test(data_norm$precipitation, data_norm$Hang, method = "pearson")
PCC_SnapLocID <- cor.test(data_norm$SnapLocID, data_norm$Hang, method = "pearson")
PCC_Snaptime <- cor.test(data_norm$Snaptime, data_norm$Hang, method = "pearson")
PCC_Temp <- cor.test(data_norm$Temp, data_norm$Hang, method = "pearson")
PCC_Turf <- cor.test(data_norm$Turf, data_norm$Hang, method = "pearson")
PCC_Wind <- cor.test(data_norm$Wind, data_norm$Hang, method = "pearson")
```

##Variance Threshold

##Caluclating Variance
```{r}
variances <- apply(data_norm, 2, var)
print(variances)

threshold <- 0.5
selected_features <- names(variances[variances >= threshold])
print(selected_features)
```
```{r}
VTDistance <- variances[selected_features[10]] 
VTGame <- variances[selected_features[15]]
VTGrass <- variances[selected_features[18]]
VTH2F <- variances[selected_features[8]]
VTOP <- variances[selected_features[7]]
VTPDate <- variances[selected_features[5]]
VTPlayerIDLS <- variances[selected_features[3]]
VTPlayerIDP <- variances[selected_features[1]]
VTPLocID <- variances[selected_features[2]]
VTPractice <- variances[selected_features[14]]
VTPrecipitation <- variances[selected_features[16]]
VTSnapLocID <- variances[selected_features[4]]
VTSnaptime <- variances[selected_features[6]]
VTTemp <- variances[selected_features[11]]
VTTurf <- variances[selected_features[10]] 

VTWind <- variances[selected_features[20]]

```


##Particle Swarm Optimization


##Pso Data
```{r}
dim(data_norm)
head(data_norm)
X <- data_norm[, -11]
head(X)
X <- X[, -9]
y <- as.factor(data_norm[,9])
```
#Packages
```{r}
install.packages("metaheuristicOpt")
install.packages("caret")
install.packages("e1071") # for SVM

library(metaheuristicOpt)
library(caret)
library(e1071)
```


##Fitness Function
```{r}
fs_fitness <- function(solution) {
  solution <- round(solution) # <-- right here, at the start

  if (sum(solution) == 0)
    return(1) # Return maximum error if no features selected

  selected <- which(solution == 1)
  X_selected <- X[, selected, drop = FALSE] # drop = FALSE to keep as data.frame/matrix even if one column

  ctrl <- trainControl(method = "cv", number = 5)

  model <- tryCatch({
    train(x = X_selected, y = y, method = "svmRadial", trControl = ctrl)
  }, error = function(e) return(NULL))

  if (is.null(model)) {
      # This handles cases where train fails (e.g., due to insufficient features for SVM)
      # Returning 1 is a good strategy to penalize failed models
      return(1)
  }

  return(1 - max(model$results$Accuracy)) # minimize error (1 - accuracy)
}


```

##Run PSO


```{r}
# Calculate the number of features (columns) in your feature matrix X
num_features <- ncol(X)

# Now your PSO call should work
res <- PSO(
  FUN = fs_fitness,
  optimType = "MIN",
  numVar = num_features, # This is now defined
  numPopulation = 20,
  maxIter = 30,
  rangeVar = matrix(c(0, 1), nrow = 2)
)
```


```{r}
ls("package:metaheuristicOpt") # Just for listing functions

best_solution <- round(res$bestVar) # round to get binary 0/1 vector
selected_features <- which(best_solution == 1)
```
```{r}
best_solution <- round(res$bestVar)  # round to get binary 0/1 vector
selected_features <- which(best_solution == 1)

```

##Defining columns for the dataframe
```{r}
column1data <- c(IG_Distance, IG_Game, IG_Grass, IG_H2F, IG_Hang, IG_Hang, IG_OP, IG_PDate, IG_PlayerIDLS, IG_PlayerIDP, IG_PLocID, IG_Practice, IG_Precipitation, IG_SnapLocID, IG_Snaptime, IG_Temp, IG_Turf, IG_Wind)
column2data <- c(MI_Distance, MI_Game, MI_Grass, MI_H2F, MI_OP, MI_PDate, MI_PlayerIDLS, MI_PlayerIDP, MI_PlocID, MI_Practice, MI_Precipitation, MI_SnapLocID, MI_Snaptime, MI_Temp, MI_Turf, MI_Wind)
column3data <- c(DistanceChi$p.value, GameChi$p.value, GrassChi$p.value, H2FChi$p.value, OPChi$p.value, DateChi$p.value, PlayerIDLSChi$p.value, PlayerIDPChi$p.value, PLocIDChi$p.value, PracticeChi$p.value, PrecipitationChi$p.value, SnapLocIDChi$p.value, SnaptimeChi$p.value, TempChi$p.value, TurfChi$p.value, WindChi$p.value) 
column4data <- c(FIDistance, FIGame, FIGrass, FIH2F, FIOP, FIPDate, FIPlayerIDLS, FIPlayerIDP, FIPLocID, FIPractice, FIPrecipitation, FISnapLocID, FISnaptime, FITemp, FITurf, FIWind)
column5data <- c(FSDistance, FSGame, FSGrass, FSH2F, FSOP, FSPDate, FSPlayerIDLS, FSPlayerIDP, FSPLocID, FSPractice,FSPrecipitation, FSSnapLocID, FSSnaptime, FSTemp, FSTurf, FSWind)
column6data <- c(PCC_Distance$estimate, PCC_Game$estimate, PCC_Grass$estimate, PCC_H2F$estimate, PCC_OP$estimate, PCC_PDate$estimate, PCC_PlayerIDLS$estimate, PCC_PlayerIDP$estimate, PCC_PLocID$estimate, PCC_Practice$estimate, PCC_Precipitation$estimate, PCC_SnapLocID$estimate, PCC_Snaptime$estimate, PCC_Temp$estimate, PCC_Turf$estimate, PCC_Wind$estimate)
column7data <- c(VTDistance, VTGame, VTGrass, VTH2F, VTOP, VTPDate, VTPlayerIDLS, VTPlayerIDP, VTPLocID, VTPractice, VTPrecipitation, VTSnapLocID, VTSnaptime, VTTemp, VTTurf, VTWind)
```

